---
title: "Fault localization experiments in Python"
author: "Mohammad Rezaalipour and Carlo A. Furia"
header-includes:
   - \usepackage[T1]{fontenc}
   - \usepackage[scaled=0.81]{beramono}
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    df_print: paged
link-citations: yes
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, cache=TRUE, cache.lazy=FALSE)
```

```{r render, eval=FALSE, include=FALSE}
## Evaluate this snippet to knit this document.
rmarkdown::render("pyfl.Rmd")
```

```{r include=FALSE}

# Assertions
library(assertthat)

# Regex and string manipulation
library(stringr)

# Powerful table manipulation
library(dplyr)
# Disable warning when summarize groups by default
options(dplyr.summarise.inform = FALSE)
library(tidyr)
## Data reshaping documentation: https://uc-r.github.io/tidyr

# Fancy, easy plots
library(ggplot2)
# More complicated plots
library(GGally)
## ggpairs extensive documentation: https://ggobi.github.io/ggally/articles/ggpairs.html

# Beautiful, readable color palettes
library(MetBrewer)

# Various effect size calculations
library(effsize)

## For determinism
my.seed <- 73544204
set.seed(my.seed)

# Colors for projects
P.PALETTE <- "Renoir"
P.COLORS <- met.brewer(P.PALETTE, n=12)

# Colors for categories
C.PALETTE <- "Archambault"
C.COLORS <- met.brewer(C.PALETTE, n=4)
```

```{r include=FALSE}

family <- "/pyfl/"
stored.values <- list()
stored.values.str <- list()

PAPER.DIR <- file.path("data")
PAPER.FNAME <- "data.tex"

STORED.NUMS <- list()
STORED.STRINGS <- list()

sanitize_ <- function(name)
{
  res <- name
  res <- gsub("&", "+", res)
  res <- gsub("%", "p", res)
  res
}

# Store `value` under `key` as numeric value.
# If `as.string` store it as string value instead.
# Use `family/group` as a hierarchical prefixes.
store_ <- function(key, value, group="", family=paste0("pyfl", DATASAMPLE), as.string=FALSE)
{
  maybe.group <- if (group == "") "" else paste0(group, "/")
  key <- sanitize_(paste0("/", family, "/", maybe.group, key))
  if (as.string)
    STORED.STRINGS[[key]] <<- seqinr::stresc(as.character(value))
  else {
    STORED.NUMS[[key]] <<- value
  }
}

# Store all content of table `df` as numeric or string values,
# under keys of the form `row.ids`:`col.ids`, using `tabname` as the group name.
#
# If `row.ids` is NULL and `df`'s row names are unique, use them as row identifiers;
# if `row.ids` is NULL and `df`'s row names are NOT unique, use row numbers as row identifiers.
# Argument `col.ids` works as `row.ids` but for column names.
#
# If `replace.probs` is TRUE and row identifiers are unique, change 
# row identifiers that denote lower bounds "|0.p" to "lower p0", 
# and row identifiers that denote upper bounds "0.p|" to "upper p0".
#
# If `vectorize` is TRUE and `df` consists of a single column, use keys 
# of the form `row.ids` (i.e., a single index instead of a pair of indexes).
store.table_ <- function(df, tabname, replace.probs=FALSE, row.ids=NULL, col.ids=NULL, vectorize=TRUE)
{
  if (is.null(row.ids)) {
    if (any(duplicated(rownames(df))))
      row.ids <- 1:nrow(df)
    else
      row.ids <- rownames(df)
  }
  if (is.null(col.ids)) {
    if (any(duplicated(colnames(df))))
      col.ids <- 1:ncol(df)
    else
      col.ids <- colnames(df)
  }
  if (replace.probs && !any(duplicated(row.ids))) {
    mo <- (row.ids %>% str_match("^[|]([0-9]+[.]?[0-9]*)$"))[, 2]
    row.ids <- ifelse(is.na(mo), row.ids, paste0("lower ", 100*as.numeric(mo)))
    mo <- (row.ids %>% str_match("^([0-9]+[.]?[0-9]*)[|]$"))[, 2]
    row.ids <- ifelse(is.na(mo), row.ids, paste0("upper ", 100*as.numeric(mo)))
  }
  for (row in 1:length(row.ids)) {
    for (col in 1:length(col.ids)) {
      if (length(col.ids) == 1 && vectorize)
        key <- paste0(row.ids[row])
      else
        key <- paste0(row.ids[row], ":", col.ids[col])
      value <- df[row, col]
      store_(key, value, group=tabname, as.string=!is.numeric(value))
    }
  }
}

# Write all numeric constants to `filename` in `out_dir`.
# If `strings`, write all string constants instead.
# If `reset`, empty the content of the variable store that has been dumped.
dump.constants_ <- function(filename, out_dir="", strings=FALSE, append=FALSE, reset=FALSE)
{
   fn <- file.path(out_dir, filename)
   if (strings)
     to.store <- STORED.STRINGS
   else
     to.store <- STORED.NUMS
   constants <- paste0("\\pgfkeyssetvalue{", names(to.store), "}{", unlist(to.store), "}")
   write_lines(constants, fn, append=append)
   if (reset) {
     if (strings)
       STORED.STRINGS <<- list()
    else
      STORED.NUMS <<- list()
   }
}

# Write all stored constants into default file
dump.all_ <- function(filename=PAPER.FNAME, out_dir=PAPER.DIR, append=TRUE, reset=TRUE)
{
  dump.constants_(filename, out_dir=out_dir, strings=FALSE, append=append)
  dump.constants_(filename, out_dir=out_dir, strings=TRUE, append=append)
}
```



# Data input

Let's start with reading the computed metrics for all projects.

```{r data input, include=TRUE}
metrics.dir <- "../metric_computation/output/"
assertthat::assert_that(dir.exists(metrics.dir))

GRANULARITIES <- factor(c("function", "statement"))
FAMILIES <- list(sbfl=c("DStar", "Ochiai", "Tarantula"), mbfl=c("Metallaxis", "Muse"), ps=c("PS"), st=c("ST"))
  
datas <- data.frame()
for (granularity in GRANULARITIES) {
  pat <- str_c("^detailed_", granularity, "_(.*)[.]csv$")
  infiles <- list.files(path=metrics.dir, pattern=pat, ignore.case=TRUE)
  for (fname in infiles) {
    technique <- str_extract(fname, regex(pat, ignore_case=TRUE), group=1)
    family <- names(which(sapply(FAMILIES, function(x) technique %in% x)))
    cur.data <- read.csv(file.path(metrics.dir, fname))
    datas <- datas %>% bind_rows(cbind(granularity=granularity, family=family, technique=technique, cur.data))
  }
}

# convert to factors relevant columns
datas$family <- factor(datas$family)
datas$technique <- factor(datas$technique)
datas$project_name <- factor(datas$project_name)

# show some data
str(datas)
```

We have a total of `r length(unique(datas$project_name))` projects, which we can classify by category:

   - Data science (`ds`): `spacy`, `keras`, `pandas`
   - Developer tools (`dev`): `cookiecutter`, `black`
   - Web tools (`web`): `sanic`, `fastapi`, `tornado`
   - Command line tools (`cl`): `httpie`, `thefuck`, `tqdm`, `youtube-dl`
   
Let's add this categorization to the data.

```{r categories}
CATEGORIES <- list(ds=c("spacy", "keras", "pandas"),
                   dev=c("cookiecutter", "black"),
                   web=c("sanic", "fastapi", "tornado"),
                   cl=c("httpie", "thefuck", "tqdm", "youtube-dl"))

categories <- sapply(datas$project_name, 
                     function(project)
                       names(which(sapply(CATEGORIES,
                                          function(categories) 
                                            project %in% categories))))
datas$category <- categories
```

# Pairwise comparisons

Let's see an example of visual and statistical comparison of two groups of experiments for the same bugs.

To make the example concrete, let's pick two groups and compare their $E_{\text{inspect}}$ scores on statement-level fault localization:

   - $S$ are the experiments done with any SBFL technique
   - $M$ are the experiments done with any MBFL technique

Since there are three experiments per bug using SBFL, but only two experiments per bug using MBFL,
we'll aggregate scores for the same bug by average.


```{r group aggregation}
data <- (datas 
         %>% filter(granularity == "statement") 
         %>% filter(family == c("sbfl", "mbfl")))
data <- (data 
         %>% group_by(family, category, project_name, bug_number) 
         %>% summarize(e_inspect=mean(e_inspect)))

S <- data %>% filter(family == "sbfl")
M <- data %>% filter(family == "mbfl")

# Match the scores in each family
joint.data <- S %>% inner_join(M, 
                               by=c("category", "project_name", "bug_number"), 
                               suffix=c(".SBFL", ".MBFL"))
```

Let's start with some visualization: a scatterplot with a point for each bug;
each point has coordinates $x, y$ where $x$ is its score in MBFL and $y$ its score in SBFL.

```{r scatterplot}
p <- ggplot(joint.data, 
            aes(x=e_inspect.MBFL, y=e_inspect.SBFL, color=category)) + geom_point(size=3) + scale_color_met_d(C.PALETTE) + theme_classic()

# x = y reference line
p <- p + geom_abline(slope=1, intercept=0)

p
```

As you can see, there are a bunch of bugs for which SBFL performs very similarly to MBFL 
(points close to the $x = y$ straight line).
However, for several other bugs, SBFL is much better (remember that lower is better for this score).

Looking at the colors, which denote the project categories, it seems data science (and possibly web frameworks) are overrepresented among the "harder" bugs on which SBFL behaves consistenly better than MBFL.

Analyzing the same data numerically, we can compute the correlation (Kendall's $\tau$) between $S$ and $M$:

```{r correlation}
corr <- cor.test(joint.data$e_inspect.SBFL, joint.data$e_inspect.MBFL, method="kendall")
corr
```

A correlation of `r corr$estimate` is not insignificant, but it is not strong either.

Finally, we may also perform a statistical test (Wilcoxon's paired test)
and compute a matching effect size (Cliff's delta).

```{r tests}
wilcox.test(joint.data$e_inspect.SBFL, joint.data$e_inspect.MBFL, paired=TRUE)

cliff.delta(joint.data$e_inspect.SBFL, joint.data$e_inspect.MBFL)
```

These statistics, for what they're worth,
seem to confirm that there is a noticeable difference in favor of SBFL.

Now, let's generalize this to a scatterplot matrix to show the relations between all possible pairs of FL families.

```{r scatterplot matrix}
FAMILIES <- c("mbfl", "ps", "st", "sbfl")

pairs_text_fn <- function(x, y, digits, corr.method, conf.level)
{
  alpha <- 1 - conf.level
  # correlation
  corObj <- stats::cor.test(x, y, method=corr.method, conf.level=conf.level)
  cor_est <- as.numeric(corObj$estimate)
  cor_txt <- formatC(cor_est, digits=digits, format="f")
  cor_sig <- corObj$p.value < alpha
  cor_abs <- abs(cor_est)
  cor_stars <- ifelse(cor_abs <= 0.3, 0,
                      ifelse(cor_abs <= 0.5, 1,
                             ifelse(cor_abs <= 0.7, 2,
                                    3)))
  cor_txt <- str_c(cor_txt, strrep("*", cor_stars))
  # Wilcoxon
  sigObj <- wilcox.test(x, y, paired=TRUE, conf.level=conf.level)
  sig_est <- as.numeric(sigObj$p.value)
  sig_txt <- formatC(sig_est, digits=digits, format="f")
  sig_sig <- sigObj$p.value < alpha
  if (sig_sig)
    sig_txt <- str_c(sig_txt, signif_stars(sigObj$p.value, point=NULL))
  # Cliff's delta
  deltaObj <- cliff.delta(x, y, conf.level=conf.level)
  delta_est <- as.numeric(deltaObj$estimate)
  delta_txt <- formatC(delta_est, digits=digits, format="f")
  delta_sig <- deltaObj$magnitude
  n_stars <- switch(as.character(delta_sig),
                    "negligible"=0,
                    "small"=1,
                    "medium"=2,
                    "large"=3)
  delta_txt <- str_c(delta_txt, strrep("*", n_stars))
  # add \n at the end of each statistic
  sapply(c(cor_txt, sig_txt, delta_txt), function(s) str_c(s, "\n"))
}

pairs.stats <- function (data, mapping, ..., 
                         digits=2, na.rm=TRUE, 
                         corr.method="kendall", conf.level=0.95)
{
    ggally_statistic(data=data, 
                     mapping=mapping, 
                     title=c("Corr", "p-value", "Effect"),
                     na.rm=na.rm,
                     justify_text="left", justify_labels="left", sep=": ",
                     title_args=list(color="black"),
                     text_fn=function(x, y) pairs_text_fn(x, y, 
                                                          digits, 
                                                          corr.method, 
                                                          conf.level))
}

points_abline <- function(data, mapping)
{
  pp <- ggplot(data=data, mapping=mapping) + geom_point()
  pp <- pp + geom_abline(intercept=0, slope=1)
  pp
}

pairwise.comparison.plot <- function(data, columns, 
                                     color_var="category", palette=NULL)
{
  pp <- ggpairs(data, columns, 
                legend=1,
                lower=list(mapping=aes(color=.data[[color_var]]),
                           continuous=wrap(points_abline)),
                diag=list(continuous=wrap("barDiag", bins=15),
                          mapping=aes(color=.data[[color_var]])),
                upper=list(continuous=wrap(pairs.stats)))
  if (!is.null(palette))
    pp <- pp + scale_color_met_d(palette) + scale_fill_met_d(palette)
  pp <- pp + theme_light()
  # change theme selectively for upper cells
  # this only works if no other theme is set in the rest of the plot
  # pp <- add_to_ggmatrix(pp, 
  #                       # this "casting" is needed in my version of GGally
  #                       ggproto(ggplot2::theme_minimal()), 
  #                       location="upper")
  return(pp)
}

data <- (datas 
         %>% filter(granularity == "statement") 
         %>% mutate(family=factor(family, levels=FAMILIES)))
data.e_inspect <- (data 
                   %>% group_by(family, category, project_name, bug_number) 
                   %>% summarize(e_inspect=mean(e_inspect)))
data.e_inspect <- data.e_inspect %>% spread(family, e_inspect)
p.e_inspect <- pairwise.comparison.plot(
  data.e_inspect, 
  columns=which(names(data.e_inspect) %in% FAMILIES),
  color_var="category", palette=C.PALETTE)
p.e_inspect
```

Now, it's easy to compute a similar plot for other metrics.

```{r pairwise other metrics}
data.time <- (data 
              %>% group_by(family, category, project_name, bug_number) 
              %>% summarize(time=mean(experiment_time_seconds)))
data.time <- data.time %>% spread(family, time)
p.time <- pairwise.comparison.plot(data.time, 
                                   columns=which(names(data.time) %in% FAMILIES),
                                   color_var="category", palette=C.PALETTE)
p.time <- p.time + theme(axis.text.x=element_text(angle=90))
p.time
```


# Dump all constants

```{r epilogue}
# dump.all_()
```