---
title: "Fault localization experiments in Python"
author: "Mohammad Rezaalipour and Carlo A. Furia"
header-includes:
   - \usepackage[T1]{fontenc}
   - \usepackage[scaled=0.81]{beramono}
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    df_print: paged
link-citations: yes
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, cache=TRUE, cache.lazy=FALSE)
```

```{r render, eval=FALSE, include=FALSE}
## Evaluate this snippet to knit this document.
rmarkdown::render("pyfl.Rmd")
```

```{r include=FALSE}

# Assertions
library(assertthat)

# Regex and string manipulation
library(stringr)
# Convenient text input output
library(readr)

# Powerful table manipulation
library(dplyr)
# Disable warning when summarize groups by default
options(dplyr.summarise.inform=FALSE)
library(tidyr)
## Data reshaping documentation: https://uc-r.github.io/tidyr

# Fancy, easy plots
library(ggplot2)
# More complicated plots
library(GGally)
## ggpairs extensive documentation: https://ggobi.github.io/ggally/articles/ggpairs.html

# Beautiful, readable color palettes
library(MetBrewer)

# Various effect size calculations
library(effsize)

# Bayesian inference
library(posterior)
library(brms)
library(rethinking)
# Calculating estimated marginal means for mixed effects models
library(emmeans)
library(marginaleffects)


## For determinism
MY.SEED <- 73544204
set.seed(MY.SEED)
# No need to thread safe RNGs
options(future.rng.onMisuse="ignore")
# Use available CPUs
options(mc.cores=parallel::detectCores())
# Set cmdstanr as backend for brms if you have it
options(brms.backend="cmdstanr")



# Colors for projects
P.PALETTE <- "Renoir"
P.COLORS <- met.brewer(P.PALETTE, n=12)

# Colors for categories
C.PALETTE <- "Archambault"
C.COLORS <- met.brewer(C.PALETTE, n=4)
```

```{r include=FALSE}

KEY.FAMILY <- "pyfl"
stored.values <- list()
stored.values.str <- list()

PAPER.DIR <- file.path("paper")
PAPER.FNAME <- "data.tex"

STORED.NUMS <- list()
STORED.STRINGS <- list()

sanitize_ <- function(name)
{
  res <- name
  res <- gsub("&", "+", res)
  res <- gsub("%", "p", res)
  res
}

# Store `value` under `key` as numeric value.
# If `as.string` store it as string value instead.
# Use `family/group` as a hierarchical prefixes.
store_ <- function(key, value, group="", family=KEY.FAMILY, as.string=FALSE)
{
  maybe.group <- if (group == "") "" else paste0(group, "/")
  key <- sanitize_(paste0("/", family, "/", maybe.group, key))
  if (as.string)
    STORED.STRINGS[[key]] <<- seqinr::stresc(as.character(value))
  else {
    STORED.NUMS[[key]] <<- value
  }
}

# Store all content of table `df` as numeric or string values,
# under keys of the form `row.ids`:`col.ids`, using `tabname` as the group name.
#
# If `row.ids` is NULL and `df`'s row names are unique, use them as row identifiers;
# if `row.ids` is NULL and `df`'s row names are NOT unique, use row numbers as row identifiers.
# Argument `col.ids` works as `row.ids` but for column names.
#
# If `replace.probs` is TRUE and row identifiers are unique, change 
# row identifiers that denote lower bounds "|0.p" to "lower p0", 
# and row identifiers that denote upper bounds "0.p|" to "upper p0".
#
# If `vectorize` is TRUE and `df` consists of a single column, use keys 
# of the form `row.ids` (i.e., a single index instead of a pair of indexes).
store.table_ <- function(df, tabname, replace.probs=FALSE, row.ids=NULL, col.ids=NULL, vectorize=TRUE)
{
  if (is.null(row.ids)) {
    if (any(duplicated(rownames(df))))
      row.ids <- 1:nrow(df)
    else
      row.ids <- rownames(df)
  }
  if (is.null(col.ids)) {
    if (any(duplicated(colnames(df))))
      col.ids <- 1:ncol(df)
    else
      col.ids <- colnames(df)
  }
  if (replace.probs && !any(duplicated(row.ids))) {
    mo <- (row.ids %>% str_match("^[|]([0-9]+[.]?[0-9]*)$"))[, 2]
    row.ids <- ifelse(is.na(mo), row.ids, paste0("lower ", 100*as.numeric(mo)))
    mo <- (row.ids %>% str_match("^([0-9]+[.]?[0-9]*)[|]$"))[, 2]
    row.ids <- ifelse(is.na(mo), row.ids, paste0("upper ", 100*as.numeric(mo)))
  }
  for (row in 1:length(row.ids)) {
    for (col in 1:length(col.ids)) {
      if (length(col.ids) == 1 && vectorize)
        key <- paste0(row.ids[row])
      else
        key <- paste0(row.ids[row], ":", col.ids[col])
      value <- df[row, col]
      store_(key, value, group=tabname, as.string=!is.numeric(value))
    }
  }
}

# Write all numeric constants to `filename` in `out_dir`.
# If `strings`, write all string constants instead.
# If `reset`, empty the content of the variable store that has been dumped.
dump.constants_ <- function(filename, out_dir="", strings=FALSE, append=FALSE, reset=FALSE)
{
   fn <- file.path(out_dir, filename)
   if (strings)
     to.store <- STORED.STRINGS
   else
     to.store <- STORED.NUMS
   constants <- paste0("\\pgfkeyssetvalue{", names(to.store), "}{", unlist(to.store), "}")
   write_lines(constants, fn, append=append)
   if (reset) {
     if (strings)
       STORED.STRINGS <<- list()
    else
      STORED.NUMS <<- list()
   }
}

# Write all stored constants into default file
dump.all_ <- function(filename=PAPER.FNAME, out_dir=PAPER.DIR, append=TRUE, reset=TRUE)
{
  dump.constants_(filename, out_dir=out_dir, strings=FALSE, append=append)
  dump.constants_(filename, out_dir=out_dir, strings=TRUE, append=append)
}
```



# Data input

Let's start with reading the computed metrics for all projects.

```{r data input, include=TRUE}
metrics.dir <- "../metric_computation/"
assertthat::assert_that(dir.exists(metrics.dir))

GRANULARITIES <- factor(c("function", "statement", "module"))
FAMILIES <- list(MBFL=c("Metallaxis", "Muse"), 
                 PS=c("PS"), 
                 ST=c("ST"),
                 SBFL=c("DStar", "Ochiai", "Tarantula"))
CATEGORIES <- list(DS=c("spacy", "keras", "pandas"),
                   DEV=c("cookiecutter", "black"),
                   WEB=c("sanic", "fastapi", "tornado"),
                   CL=c("httpie", "thefuck", "tqdm", "youtubedl"))
  
datas <- data.frame()
for (granularity in GRANULARITIES) {
  csvpath <- file.path(metrics.dir, 
                       str_c("output_fauxpy_", granularity),
                       str_c("fauxpy_", granularity, "_detailed.csv"))
  cur.data <- read.csv(csvpath)
  datas <- datas %>% bind_rows(cur.data)
}

# rename columns
names(datas)[names(datas) == "project_name"] <- "project"
names(datas)[names(datas) == "e_inspect"] <- "einspect"
names(datas)[names(datas) == "experiment_time_seconds"] <- "time"
names(datas)[names(datas) == "exam_score"] <- "exam"
names(datas)[names(datas) == "cumulative_distance"] <- "cdist"
names(datas)[names(datas) == "sv_comp_overall_score"] <- "svcomp"
names(datas)[names(datas) == "mutable_bug"] <- "mutable"
names(datas)[names(datas) 
             == "percentage_of_mutants_on_ground_truth"] <- "mutability"

# rename to avoid - character
datas$project[datas$project == "youtube-dl"] <- "youtubedl"

# use booleans where appropriate
datas$crashing <- as.logical(datas$crashing)
datas$predicate <- as.logical(datas$predicate)
datas$mutable <- as.logical(datas$mutable)

# convert to factors
datas$project <- factor(datas$project)
datas$granularity <- factor(datas$granularity)
levels(datas$granularity) <- GRANULARITIES
datas$technique <- factor(datas$technique)
# add time in minutes
datas$minutes <- datas$time / 60

TECHNIQUES <- unique(datas$technique)
PROJECTS <- unique(datas$project)

#' Given a technique, which family it belongs to
which.family <- function(t) names(FAMILIES)[sapply(FAMILIES, function(x) t %in% x)]

# Derive each technique's family
datas$family <- sapply(datas$technique, which.family)
datas$family <- factor(datas$family)
# Force order of families as in names(FAMILIES)
datas <- (datas %>% mutate(family=factor(family, levels=names(FAMILIES))))

#' Given a project, which category it belongs to
which.category <- function(p) names(CATEGORIES)[sapply(CATEGORIES, 
                                                       function(x) p %in% x)]

# Derive each project's category
datas$category <- sapply(datas$project, which.category)
datas$category <- factor(datas$category)

# Assign unique identifier to every bug
bugs <- str_c(datas$project, datas$bug_number)
bugs <- factor(bugs)
datas$bugid <- bugs
BUGS <- unique(datas$bug)

#' Given a bug unique identifier, which category of programs it belongs to
which.projcategory <- function(t) 
  names(CATEGORIES)[sapply(CATEGORIES, 
                           function(x) any(str_detect(t, x)))]

# show some data
str(datas)
```

We have a total of `r length(unique(datas$project))` projects, which we can classify by category:

   - Data science (`ds`): `spacy`, `keras`, `pandas`
   - Developer tools (`dev`): `cookiecutter`, `black`
   - Web tools (`web`): `sanic`, `fastapi`, `tornado`
   - Command line tools (`cl`): `httpie`, `thefuck`, `tqdm`, `youtubedl`
   

# Pairwise comparisons

Let's see an example of visual and statistical comparison of two groups of experiments for the same bugs.

To make the example concrete, let's pick two groups and compare their $E_{\text{inspect}}$ scores on statement-level fault localization:

   - $S$ are the experiments done with any SBFL technique
   - $M$ are the experiments done with any MBFL technique

Since there are three experiments per bug using SBFL, 
but only two experiments per bug using MBFL,
we'll aggregate scores for the same bug by average.


```{r group aggregation}
statement.by.family.einspect <- (
  datas 
  %>% filter(granularity == "statement")
  %>% group_by(family, category, project, bug_number)
  %>% summarize(einspect=mean(einspect))
  %>% spread(family, einspect)
)

statement.einspect.upper <- max(unlist(
  statement.by.family.einspect[, 4:ncol(statement.by.family.einspect)])
  )

S <- statement.by.family.einspect$SBFL
M <- statement.by.family.einspect$MBFL
```

Let's start with some visualization: a scatterplot with a point for each bug;
each point has coordinates $x, y$ where $x$ is its score in MBFL and $y$ its score in SBFL.

```{r scatterplot}
p <- ggplot(statement.by.family.einspect, 
            aes(x=MBFL, y=SBFL, color=category)) + geom_point(size=3)
p <- (p + scale_color_met_d(C.PALETTE) 
      + scale_x_continuous(limits=c(0, statement.einspect.upper))
      + scale_y_continuous(limits=c(0, statement.einspect.upper)))
p <- p + theme_classic()

# x = y reference line
p <- p + geom_abline(slope=1, intercept=0)

p
```

As you can see, 
there are a bulk of bugs for which SBFL performs very similarly to MBFL 
(points close to the $x = y$ straight line).
However, for several other bugs, SBFL is much better (remember that lower is better for this score).

Looking at the colors, we notice that several bugs in the CL (and possibly DS) category are overrepresented among the "harder" bugs on which SBFL behaves much better than MBFL.

Analyzing the same data numerically, we can compute the correlation (Kendall's $\tau$) between $S$ and $M$:

```{r correlation}
(corr <- cor.test(S, M, method="kendall"))
```

A correlation of `r corr$estimate` is not very strong, but clearly defined.

Finally, we may also perform a statistical test (Wilcoxon's paired test)
and compute a matching effect size (Cliff's delta).

```{r tests}
wilcox.test(S, M, paired=TRUE)

cliff.delta(S, M)
```

Cliff's delta, in particular, roughly measures how often the value in one set are larger than the value in the other set.
Thus, the given value means that SBFL's $E_{\text{inspect}}$ score is smaller than MBFL's roughly in 19\% of the cases.


These statistics, for what they're worth,
seem to confirm that there is a noticeable difference in favor of SBFL.

Now, let's generalize this to a scatterplot matrix to show the relations between all possible pairs of FL families.

First, we define a bunch of helper functions.

```{r scatterplot matrix funs, echo=FALSE, message=FALSE, results='hide', fig.keep='all'}
#' Compute statistics comparing data `x` and `y`.
#'
#' The function returns a vector of strings, each expressing a statistics and
#' some kind of strength level with stars.
#'
#' @param x   A numeric vector with the first dataset.
#' @param y   A numeric vector with the second dataset.
#' @param digits   The number of decimal digits of the statistics.
#' @param corr.method   The correlation method to be used (see `cor.test`).
#' @param conf.level   The confidence level used for the significance tests.
#' @param info  If not NULL (the default), a list with components `granularity`, 
#'   `metric`, `g1`, `g2`, used to store key constants of the various stats.
#'
#' @return  Returns a vector of strings with correlation, p-value of 
#'          a Wilcoxon paired test, and Cliff's delta.
#'
pairs_text_fn <- function(x, y, digits, corr.method, conf.level,
                          info=NULL)
{
  alpha <- 1 - conf.level
  # correlation
  corObj <- stats::cor.test(x, y, method=corr.method, conf.level=conf.level)
  cor_est <- as.numeric(corObj$estimate)
  cor_txt <- formatC(cor_est, digits=digits, format="f")
  cor_sig <- corObj$p.value < alpha
  cor_abs <- abs(cor_est)
  cor_stars <- ifelse(cor_abs <= 0.3, 0,
                      ifelse(cor_abs <= 0.5, 1,
                             ifelse(cor_abs <= 0.7, 2,
                                    3)))
  cor_txt <- str_c(cor_txt, strrep("*", cor_stars))
  # Wilcoxon
  sigObj <- wilcox.test(x, y, paired=TRUE, conf.level=conf.level)
  sig_est <- as.numeric(sigObj$p.value)
  sig_txt <- formatC(sig_est, digits=digits, format="f")
  sig_sig <- sigObj$p.value < alpha
  if (sig_sig)
    sig_txt <- str_c(sig_txt, signif_stars(sigObj$p.value, point=NULL))
  # Cliff's delta
  deltaObj <- cliff.delta(x, y, conf.level=conf.level)
  delta_est <- as.numeric(deltaObj$estimate)
  delta_txt <- formatC(delta_est, digits=digits, format="f")
  delta_sig <- deltaObj$magnitude
  n_stars <- switch(as.character(delta_sig),
                    "negligible"=0,
                    "small"=1,
                    "medium"=2,
                    "large"=3)
  delta_txt <- str_c(delta_txt, strrep("*", n_stars))
  # add \n at the end of each statistic
  res <- c(sapply(c(cor_txt, sig_txt), function(s) str_c(s, "\n")), delta_txt)
  if (!is.null(info)) {
    grp <- str_c(info$granularity, "/", info$metric, "/", info$g1,":", info$g2)
    store_("corrTau", cor_est, grp)
    store_("WilcoxonP", sig_est, grp)
    store_("CliffD", delta_est, grp)
    return(grp)
  }
  res
}

#' Helper function wrapping `pairs_text_fn` in a call to 
#' `GGally::ggally_statistics`, providing sensible defaults.
pairs.stats <- function (data, mapping, ..., 
                         digits=2, na.rm=TRUE, 
                         corr.method="kendall", conf.level=0.95)
{
    ggally_statistic(data=data, 
                     mapping=mapping, 
                     title=c("Corr", "p-value", "Effect"),
                     na.rm=na.rm,
                     justify_text="left", justify_labels="left", sep=": ",
                     title_args=list(color="black"),
                     text_fn=function(x, y) pairs_text_fn(x, y, 
                                                          digits, 
                                                          corr.method, 
                                                          conf.level))
}

#' Helper function to add a line x = y to a ggpairs plot.
points_abline <- function(data, mapping, ...)
{
  pp <- ggplot(data=data, mapping=mapping) + geom_point()
  if (hasArg("lower_limits")) {
    extra.params <- list(...)
    pp <- (pp 
           + scale_x_continuous(limits=extra.params$lower_limits) 
           + scale_y_continuous(limits=extra.params$lower_limits))
  }
  pp <- pp + geom_abline(intercept=0, slope=1)
  pp
}

#' Helper function to set the scale to a ggpairs bar plot.
bar_scale <- function(data, mapping)
{
  pp <- ggally_barDiag(data, mapping, bins=15)
  # pp <- pp + scale_x_continuous(limits=statement.einspect.limits)
  pp
}


#' Pairwise comparison plot with scatterplots on lower, bar charts on diagonal, 
#' and statistics on upper.
#' 
#' @param data  The dataset used for plotting.
#' @param columns  The indexes of columns to be compared pairwise.
#' @param color_var  The column used to group the plotted data, as a string.
#' @param palette  The name of the MetBrewer palette used to color groups.
#'
#' @return  Returns a ggplot2 plot object. 
pairwise.comparison.plot <- function(data, columns, 
                                     color_var="category", palette=NULL, ...)
{
  pp <- ggpairs(data, columns, 
                legend=1,
                lower=list(mapping=aes(color=.data[[color_var]]),
                           continuous=wrap(points_abline, ...)),
                diag=list(continuous=wrap(bar_scale),
                          mapping=aes(color=.data[[color_var]])),
                upper=list(continuous=wrap(pairs.stats)))
  if (!is.null(palette))
    pp <- pp + scale_color_met_d(palette) + scale_fill_met_d(palette)
  pp <- pp + theme_light() + theme(
    legend.title=element_blank(),
    legend.position="bottom",
    legend.direction="horizontal"
    # aspect.ratio=3/6
  )
  # change theme selectively for upper cells
  # this only works if no other theme is set in the rest of the plot
  # pp <- add_to_ggmatrix(pp, 
  #                       # this "casting" is needed in my version of GGally
  #                       ggproto(ggplot2::theme_minimal()), 
  #                       location="upper")
  return(pp)
}
```

Then, we use them to generate plots for $E$.


```{r scatterplot einspect, echo=FALSE, message=FALSE, results='hide', fig.keep='all'}
statement.einspect.limits <- c(0, statement.einspect.upper)

p.einspect <- pairwise.comparison.plot(
  statement.by.family.einspect, 
  columns=which(names(statement.by.family.einspect) %in% names(FAMILIES)),
  color_var="category", palette=C.PALETTE,
  lower_limits=statement.einspect.limits)
p.einspect

f.p.einspect <- file.path(PAPER.DIR, "statement-family-einspect.pdf") 
ggsave(f.p.einspect, 
       plot=p.einspect, 
       device="pdf")
knitr::plot_crop(f.p.einspect, quiet=TRUE)

all.pairs <- statement.by.family.einspect[, 
                                          (colnames(statement.by.family.einspect) 
                                           %in% names(FAMILIES))]
# generate all pairs
all.pairs <- combn(all.pairs, 2, simplify=FALSE)

# also save statistics as constants
x_ <- lapply(all.pairs, 
             function(p) 
               pairs_text_fn(unlist(p[1]), unlist(p[2]), 2, "kendall", 0.95,
                             info=list(
                               granularity="statement", 
                               metric="einspect", 
                               g1=colnames(p)[1], 
                               g2=colnames(p)[2])))
```

Now, it's easy to compute a similar plot for other metrics.
For example, running time (in minutes):

```{r scatterplot time, echo=FALSE, message=FALSE, results='hide', fig.keep='all'}

statement.by.family.time <- (
  datas 
  %>% filter(granularity == "statement")
  %>% group_by(family, category, project, bug_number)
  %>% summarize(minutes=mean(minutes))
  %>% spread(family, minutes)
)

statement.time.upper <- max(unlist(
  statement.by.family.time[, 4:ncol(statement.by.family.time)])
  )

statement.time.limits <- c(0, statement.time.upper)

p.time <- pairwise.comparison.plot(
  statement.by.family.time, 
  columns=which(names(statement.by.family.time) %in% names(FAMILIES)),
  color_var="category", palette=C.PALETTE,
  lower_limits=statement.time.limits)
# p.time <- p.time + theme(axis.text.x=element_text(angle=90))
p.time

f.p.time <- file.path(PAPER.DIR, "statement-family-time.pdf") 
ggsave(f.p.time, 
       plot=p.time, 
       device="pdf")
knitr::plot_crop(f.p.time, quiet=TRUE)

all.pairs <- statement.by.family.time[, 
                                      (colnames(statement.by.family.time) 
                                       %in% names(FAMILIES))]
# generate all pairs
all.pairs <- combn(all.pairs, 2, simplify=FALSE)

# also save statistics as constants
x_ <- lapply(all.pairs, 
             function(p) 
               pairs_text_fn(unlist(p[1]), unlist(p[2]), 2, "kendall", 0.95,
                             info=list(
                               granularity="statement", 
                               metric="time", 
                               g1=colnames(p)[1], 
                               g2=colnames(p)[2])))
```


# Regression models

Let's build a simple multivariate regression model, 
where we predict `einspect` and `time` from bug and technique.

First, we standardize the predictors, so that it's much easier to set sensible priors.

```{r bug identifiers}
by.statement <- (datas %>% filter(granularity=="statement"))

by.statement$timeS <- ((by.statement$time - mean(by.statement$time))
                       / sd(by.statement$time))
by.statement$einspectS <- ((by.statement$einspect - mean(by.statement$einspect))
                           / sd(by.statement$einspect))
```

## Model $m_1$

Here's a basic regression model, where the only unusual aspects
are that it's multivariate, and log-transforms the mean (since both outcome variables must be nonnegative).

```{r m1 formula}
eq.m1 <- brmsformula(
  mvbind(einspectS, timeS) ~ 0 + family + category,
  family=brmsfamily("gaussian", link="log")
) + set_rescor(TRUE)

pp1.check <- get_prior(eq.m1, data=by.statement)

pp1 <- c(
  set_prior("normal(0, 1.0)", class="b", resp=c("einspectS", "timeS")),
  set_prior("weibull(2, 1)", class="sigma", resp=c("einspectS", "timeS"))
)
```


## Fitting $m_1$

Let's do the usual checks to make sure that everything is fine with the fitting.

Prior checks, confirming that the sampled priors span a wide range of values, 
amply including the data.

```{r m1 priors check, warning=FALSE, message=FALSE, error=FALSE, results='hide'}
m1.priors <- brm(eq.m1,
                 prior=pp1,
                 data=by.statement,
                 chains=1, iter=1000,
                 sample_prior="only",
                 refresh=0, 
                 control=list(adapt_delta=0.95),
                 seed=MY.SEED)

(prior1.check.einspect <- (pp_check(m1.priors, draw_ids=1:100, resp="einspectS") 
                           + scale_x_log10()))
(prior1.check.time <- (pp_check(m1.priors, draw_ids=1:100, resp="timeS") 
                       + scale_x_log10()))
```

Now we fit the actual model.

```{r m1 fit}
m1 <- brm(eq.m1, data=by.statement, prior=pp1, refresh=0,
          control=list(adapt_delta=0.95), seed=MY.SEED)
```

Next, we check the usual diagnostics:

   - No (or at most a few) divergent transitions
   - $\widehat{R}$ ratio below $1.01$
   - Effective sample size (ESS), as a ratio of the total sample size, at least 10%

```{r m1 diagnostics}

# Divergent transitions
np <- nuts_params(m1)
print(sum(subset(np, Parameter == "divergent__")$Value))

# \hat{R} ratio
print(max(brms::rhat(m1), na.rm = TRUE))

# ESS
min(neff_ratio(m1), na.rm = TRUE)
```

Finally, we check the posteriors, to ensure that we have a decent approximation of the data.

```{r m1 posterior checks, warning=FALSE, message=FALSE}
print(pp_check(m1, ndraws=100, resp="einspectS") + scale_x_log10())
print(pp_check(m1, ndraws=100, resp="timeS") + scale_x_log10())
```

As you can see, the simulated posteriors are acceptable given that 
the data is complex, whereas the model is quite simple.


## Analyzing $m_1$


```{r m1 summary}
summary(m1)
```
What's noticeable here is some non-negligible residual correlation
between the two outcomes `einspect` and `time`, which means that there
is a dependency between these two variables that is not captured by the model.

Let's set up some functions to analyze the posterior samples of `m1` (and similar models).

```{r m1 difference analysis defs, warning=FALSE}
PROBABILITIES <- c(0.5, 0.7, 0.9, 0.95, 0.99)

interval.results <- function(fit, response, which.group, predictor)
{
  # posterior draws
  post <- posterior::as_draws_df(fit)
  # get columns of coefficients corresponding to `response` and `predictor`
  post <- post[, grep(str_c("^b_", response, "_", predictor), colnames(post))]
  # overall mean of posterior draws
  grand.mean <- mean(unlist(post))
  # difference to mean
  diffs <- apply(post, 2, function(x) x - grand.mean)
  # keep only the predictor's names
  colnames(diffs) <- str_extract(colnames(diffs), 
                                 str_c("^b_", response, "_", predictor, "(.*)$"),
                                 group=1)
  # reshape
  diffs <- data.frame(diffs) %>% gather(predictor, response)
  # add group column
  diffs$group <- sapply(diffs$predictor, which.group)
  # compute HPDIs
  intervals <- (diffs
                %>% group_by(group) 
                %>% reframe(response=HPDI(response, prob=rev(PROBABILITIES)), 
                            prob=names(HPDI(response, prob=rev(PROBABILITIES)))))
  # reshape back and sort in inclusion interval order
  intervals <- intervals %>% spread(group, response)
  # sort in inclusion interval order
  nrow.2 <- round(nrow(intervals)/2)
  res <- rbind(intervals[1:nrow.2, ], intervals[nrow(intervals):(nrow.2+1), ])
  rnms <- res$prob
  res <- as.data.frame(res[, 2:ncol(res)])
  rownames(res) <- rnms
  list(ints=res, est=NULL)
}
```

Let's use these functions to first analyze the effects per *family*
of FL techniques.

```{r m1 per family, warning=FALSE}
(res.m1.einspect.family <- interval.results(m1, "einspectS", 
                                            identity, "family"))
(res.m1.time.family <- interval.results(m1, "timeS", 
                                        identity, "family"))
```
It's clear that for both outcomes, `e_inspect` and `time`,
there are strong differences (with high probability) 
in the contribution over the mean from different families of techniques.
The only exception is for MBFL and `e_inspect`, where the effect is a wash
(probably because the two techniques work fairly differently).

Looking at the effects by category of project does not
yeld as strong differences, but we can see that `ds` projects tend to be associated with worse `e_inspect`, and that `web` projects are assocaited with 
shorter running times.

```{r m1 per category, warning=FALSE}
(res.m1.einspect.category <- interval.results(m1, "einspectS", 
                                              identity, "category"))
(res.m1.time.category <- interval.results(m1, "timeS", 
                                          identity, "category"))
```

## Model $m_2$

Let's make the model more sophisticated, with varying effects,
and modeling these effects as [possibly correlated](https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html) 
(which makes sense,
since we have two model parts)

```{r m2 formula}
eq.m2 <- brmsformula(
  mvbind(einspectS, timeS) ~ 1 + (1|p|family) + (1|q|category),
  family=brmsfamily("gaussian", link="log")
) + set_rescor(TRUE)

pp2.check <- get_prior(eq.m2, data=by.statement)

pp2 <- c(
  set_prior("normal(0, 1.0)", class="Intercept", resp=c("einspectS", "timeS")),
  set_prior("weibull(2, 0.3)", class="sd", coef="Intercept", 
            group="family", resp=c("einspectS", "timeS")),
  set_prior("weibull(2, 0.3)", class="sd", coef="Intercept", 
            group="category", resp=c("einspectS", "timeS")),
  set_prior("gamma(0.01, 0.01)", class="sigma", resp=c("einspectS", "timeS"))
)
```


## Fitting $m_2$

Let's fit $m_2$ and check the fit.

Prior checks:

```{r m2 priors check, warning=FALSE, message=FALSE, error=FALSE, results='hide'}
m2.priors <- brm(eq.m2,
                 prior=pp2,
                 data=by.statement,
                 chains=1, iter=1000,
                 sample_prior="only",
                 refresh=0, 
                 control=list(adapt_delta=0.95),
                 seed=MY.SEED)

(pp_check(m2.priors, draw_ids=1:100, resp="einspectS") + scale_x_log10())
(pp_check(m2.priors, draw_ids=1:100, resp="timeS") + scale_x_log10())
```

We fit model $m_2$.

```{r m2 fit}
m2 <- brm(eq.m2, data=by.statement, prior=pp2, 
          refresh=0,
          control=list(adapt_delta=0.95), 
          seed=MY.SEED)
```

Diagnostics:

```{r m2 diagnostics}

# Divergent transitions
np <- nuts_params(m2)
print(sum(subset(np, Parameter == "divergent__")$Value))

# \hat{R} ratio
print(max(brms::rhat(m2), na.rm = TRUE))

# ESS
min(neff_ratio(m2), na.rm = TRUE)
```

Posterior checks:

```{r m2 posterior checks, warning=FALSE, message=FALSE}
print(pp_check(m2, ndraws=100, resp="einspectS") + scale_x_log10())
print(pp_check(m2, ndraws=100, resp="timeS") + scale_x_log10())
```

We don't notice a clear improvement compared to $m_1$.
Let's compare the two models using LOO.

```{r model comparison, warning=FALSE}
m1 <- add_criterion(m1, criterion = "loo")
m2 <- add_criterion(m2, criterion = "loo")

(lc.m1m2 <- loo(m1, m2, moment_match = TRUE))
```

$m_1$'s score is around 2.4 standard deviations worse than $m_2$'s,
which is a non-negligible (albeit not strong) difference in favor of $m_2$
in terms of predictive capabilities.

## Analyzing $m_2$


```{r m2 summary}
summary(m2)
```
By `family`, there is a slight inverse correlation between
the two outcomes `einspect` and `time`; this correlation disappears 
if we look at the `category` terms. 
The residual correlation is about the same as in `m_1`.

Let's now perform an effects analysis 
on the fitted coefficients of `m2`.
First we introduce a summary function.

```{r m2 difference analysis defs, warning=FALSE}
interval.results.ranef <- function(fit, response, predictor, probs, 
                                   multivariate=TRUE)
{
  # probability intervals and their labels
  ps.low <- rev(1-probs)/2
  ps.high <- rev(1-ps.low)
  ps <- rev(c(ps.high, ps.low))
  cnams <- c(paste("|", probs, sep=""), paste(rev(probs), "|", sep=""))
  # random effects
  post <- ranef(fit, probs=ps)
  if (multivariate) {
    # keep only predictor/response data
    post <- post[[predictor]]
    est <- post[, "Estimate", 1]
    post <- post[, , grep(str_c("^", response), dimnames(post)[[3]])]
  } else {
    post <- post[[predictor[1]]]
    post <- post[, , grep(str_c("^", predictor[2]), dimnames(post)[[3]])]
    est <- post[, grep("Estimate", colnames(post))]
  }
  if (dims(post) > 2 && !("matrix" %in% class(post)))
    post <- post[, , 1]
  res <- as.data.frame(post)
  res <- res[, grep("^Q", colnames(res))]
  colnames(res) <- cnams
  res <- t(res)
  
  list(ints=res, est=est)
}
```

Then we use the functions to analyze the effects
of the FL techniques.

```{r m2 per family, warning=FALSE}
(res.m2.einspect.family <- interval.results.ranef(
  m2, "einspectS",  "family", PROBABILITIES
 ))
(res.m2.time.family <- interval.results.ranef(
  m2, "timeS",  "family", PROBABILITIES
 ))
```
The results for `einspect` are clearer than with the simpler model,
whereas the relation between running time and family `ps` weakens.

Let's see what happens for the bug/category of projects.

```{r m2 per category, warning=FALSE}
(res.m2.einspect.category <- interval.results.ranef(
  m2, "einspectS", "category", PROBABILITIES)
 )
(res.m2.time.category <- interval.results.ranef(
  m2, "timeS", "category", PROBABILITIES)
  )
```

## Model $m_3$

```{r m3 formula}
eq.m3 <- brmsformula(
  mvbind(einspectS, timeS) ~ 
    0 + family + category + (0 + family|r|category),
  family=brmsfamily("gaussian", link="log")
) + set_rescor(TRUE)

pp3.check <- get_prior(eq.m3, data=by.statement)

pp3 <- c(
  set_prior("normal(0, 1.0)", class="b", resp=c("einspectS", "timeS")),
  set_prior("gamma(0.01, 0.01)", class="sigma", resp=c("einspectS", "timeS")),
  set_prior("lkj(1)", class="cor"),
  set_prior("weibull(2, 0.3)", class="sd", resp=c("einspectS", "timeS"))
)
```


## Fitting $m_3$

Let's fit $m_3$ and check the fit.

Prior checks:

```{r m3 priors check, warning=FALSE, message=FALSE, error=FALSE, results='hide'}
m3.priors <- brm(eq.m3,
                 prior=pp3,
                 data=by.statement,
                 chains=1, iter=1000,
                 sample_prior="only",
                 refresh=0, 
                 control=list(adapt_delta=0.95),
                 seed=MY.SEED)

(pp_check(m3.priors, draw_ids=1:100, resp="einspectS") + scale_x_log10())
(pp_check(m3.priors, draw_ids=1:100, resp="timeS") + scale_x_log10())
```

We fit model $m_3$.

```{r m3 fit}
m3 <- brm(eq.m3, data=by.statement, prior=pp3, 
          refresh=0,
          control=list(adapt_delta=0.95),
          seed=MY.SEED)
```

Diagnostics:

```{r m3 diagnostics}
# Divergent transitions
np <- nuts_params(m3)
print(sum(subset(np, Parameter == "divergent__")$Value))

# \hat{R} ratio
print(max(brms::rhat(m3), na.rm = TRUE))

# ESS
min(neff_ratio(m3), na.rm = TRUE)
```
Posterior checks:

```{r m3 posterior checks, warning=FALSE, message=FALSE}
print(pp_check(m3, ndraws=100, resp="einspectS") + scale_x_log10())
print(pp_check(m3, ndraws=100, resp="timeS") + scale_x_log10())
```

In line with what seen before, possibly a bit better.

Model comparison:

```{r model comparison, warning=FALSE}
m3 <- add_criterion(m3, criterion = "loo")

(lc.m1m2m3 <- loo(m1, m2, m3, moment_match = TRUE))
```

$m_2$'s score is only one standard deviation worse than $m_2$'s.
This is modest improvement, and hence it may not be worth to 
go with the much more complex model $m_3$.


## Analyzing $m_3$


```{r m2 summary}
summary(m3)
```

Instead of considering the fixed and varying effects of
$m_3$, we can estimate the marginal means for each family of FL techniques.

```{r m3 per family, warning=FALSE}
cnams.lows <- paste("|", PROBABILITIES, sep="")
cnams.highs <- paste(PROBABILITIES, "|", sep="")
cnams <- c(cnams.lows, cnams.highs)[order(
  c(seq_along(cnams.lows), seq_along(cnams.highs)
    ))]

# Very similar results with: 
# marginaleffects::avg_predictions(m3, by="family", conf_level=p)
ems <- lapply(PROBABILITIES, 
              function(p) as.data.frame(emmeans(m3, "family", level=p)))
ests <- ems[[1]][, 1:2]
ints <- lapply(ems, function(df) df[, 3:ncol(df)])
# merge dataframes in list
ints <- do.call("cbind", ints)
colnames(ints) <- cnams
# resort columns according to prob order
idxs <- c(seq(1, length(cnams), by=2), rev(1 + seq(1, length(cnams), by=2)))
ints <- ints[, idxs]
# add names of families
rownames(ints) <- ests$family
# transpose
ints <- t(ints)
# get vector of estimates
est <- ests$emmean
names(est) <- ests$family
res.m3.einspect.family <- list(ints=ints, est=est)
```


## Model $m_4$

Let's now add predictors to $m_2$, so as to study 
any effect of the kinds of bugs:

   - `predicate` is a Boolean value that identifies predicate-related bugs
   - `crashing` is a Boolean value that identifies crashing bugs
   - `mutability` is a nonnegative score that denotes the percentage of
     mutants that mutate a line in a bug's ground truth
   - `mutable` is a Boolean that identifies the bugs with a 
     positive mutability score
     
Since `mutability`/`mutable` are likely affecting `category`
and `einspect`, it makes sense to add the predictor,
so as to close the possible backdoor path $\textrm{category} \leftarrow \textrm{mutable} \rightarrow \textrm{einspect}$.
     
```{r m4 formula}
eq.m4 <- brmsformula(
  mvbind(einspectS, timeS) ~ 1 + (1|p|family) + (1|q|category) 
  + predicate*family + crashing*family + mutable*family,
  family=brmsfamily("gaussian", link="log")
) + set_rescor(TRUE)

pp4.check <- get_prior(eq.m4, data=by.statement)

pp4 <- c(
  set_prior("normal(0, 1.0)", class="Intercept", resp=c("einspectS", "timeS")),
  set_prior("normal(0, 1.0)", class="b", resp=c("einspectS", "timeS")),
  set_prior("weibull(2, 0.3)", class="sd", coef="Intercept", 
            group="family", resp=c("einspectS", "timeS")),
  set_prior("weibull(2, 0.3)", class="sd", coef="Intercept", 
            group="category", resp=c("einspectS", "timeS")),
  set_prior("gamma(0.01, 0.01)", class="sigma", resp=c("einspectS", "timeS"))
)
```


## Fitting $m_4$

Prior checks:

```{r m4 priors check, warning=FALSE, message=FALSE, error=FALSE, results='hide'}
m4.priors <- brm(eq.m4,
                 prior=pp4,
                 data=by.statement,
                 chains=1, iter=1000,
                 sample_prior="only",
                 refresh=0, 
                 control=list(adapt_delta=0.95),
                 seed=MY.SEED)

(pp_check(m4.priors, draw_ids=1:100, resp="einspectS") + scale_x_log10())
(pp_check(m4.priors, draw_ids=1:100, resp="timeS") + scale_x_log10())
```

We fit model $m_4$.

```{r m2 fit}
m4 <- brm(eq.m4, data=by.statement, prior=pp4, 
          refresh=0,
          control=list(adapt_delta=0.95), 
          seed=MY.SEED)
```

Diagnostics:

```{r m4 diagnostics}

# Divergent transitions
np <- nuts_params(m4)
print(sum(subset(np, Parameter == "divergent__")$Value))

# \hat{R} ratio
print(max(brms::rhat(m4), na.rm = TRUE))

# ESS
min(neff_ratio(m4), na.rm = TRUE)
```

Posterior checks:

```{r m4 posterior checks, warning=FALSE, message=FALSE}
print(pp_check(m4, ndraws=100, resp="einspectS") + scale_x_log10())
print(pp_check(m4, ndraws=100, resp="timeS") + scale_x_log10())
```

Let's compare the models using LOO.

```{r model comparison, warning=FALSE}
m4 <- add_criterion(m4, criterion = "loo")

(lc.m1m2m3m4 <- loo(m1, m2, m3, m4, moment_match = TRUE))
```

Nearly 2 standard deviations of improvement is not bad.


## Analyzing $m_4$


```{r m4 summary}
summary(m4)
```
Let's now perform an effects analysis 
on the fitted coefficients of `m4`.

Specifically, we look at the (fixed) effects
of the families associated with certain categories of bugs.
We focus on `einspect` only, since the kind of bugs is not 
meaningfully associated with the running time.

```{r m4 analysis, warning=FALSE}
interval.results.interacts <- function(fit, response, predictor, probs, 
                                       bool=TRUE, multivariate=TRUE)
{
  ps.low <- rev(1-probs)/2
  ps.high <- rev(1-ps.low)
  ps <- rev(c(ps.high, ps.low))
  cnams <- c(paste("|", probs, sep=""), paste(rev(probs), "|", sep=""))
  post <- fixef(fit, probs=ps)
  if (multivariate)
    pat <- str_c("^", response, "_.*", predictor, if (bool) "TRUE" else "")
  else
    pat <- str_c("^.*", predictor, if (bool) "TRUE" else "")
  post <- post[grep(pat, rownames(post)), ]
  ints <- post[, 3:ncol(post)]
  ests <- post[, 1]
  labs <- str_c(predictor, " ", names(FAMILIES))
  names(ests) <- labs
  colnames(ints) <- cnams
  rownames(ints) <- labs
  ints <- t(ints)
  res <- list(ints=ints, est=ests)
  res
}

(res.m4.einspect.crashing <- interval.results.interacts(m4, "einspectS", 
                                                        "crashing", PROBABILITIES))
(res.m4.einspect.predicate <- interval.results.interacts(m4, "einspectS", 
                                                         "predicate",
                                                         PROBABILITIES))
(res.m4.einspect.mutable <- interval.results.interacts(m4, "einspectS", 
                                                       "mutable", PROBABILITIES))
```
So, crashing bugs are indeed easier for ST.
In contrast, predicate-related bugs do not seem to be simpler
for PS, although they are a bit harder for ST.

For the mutability bugs, we probably have to include 
in the model a clearer dependency on the `mutability`
rather than just the boolean indicator `mutable`.


## Model $m_5$

A simple way would be to introduce an interaction `mutability`$\times$`family`.

```{r m4 formula}
eq.m5 <- brmsformula(
  mvbind(einspectS, timeS) ~ 1 + (1|p|family) + (1|q|category) 
  + predicate*family + crashing*family + mutability*family,
  family=brmsfamily("gaussian", link="log")
) + set_rescor(TRUE)

pp5.check <- get_prior(eq.m5, data=by.statement)

pp5 <- c(
  set_prior("normal(0, 1.0)", class="Intercept", resp=c("einspectS", "timeS")),
  set_prior("normal(0, 1.0)", class="b", resp=c("einspectS", "timeS")),
  set_prior("weibull(2, 0.3)", class="sd", coef="Intercept", 
            group="family", resp=c("einspectS", "timeS")),
  set_prior("weibull(2, 0.3)", class="sd", coef="Intercept", 
            group="category", resp=c("einspectS", "timeS")),
  set_prior("gamma(0.01, 0.01)", class="sigma", resp=c("einspectS", "timeS"))
)
```

We could get passable prior checks, but let's cut to the chase
and fit model $m_5$.

```{r m5 fit}
m5 <- brm(eq.m5, data=by.statement, prior=pp5, 
          refresh=0,
          control=list(adapt_delta=0.95), 
          seed=MY.SEED)
```

The first thing that we notice is that one of the chains
went awry and spinned for much longer than all the other chains.
In addition, we got a number of scary warnings.
This points to some region of the posterior that could not be sampled
effectively.

Let's see the diagnostics:

```{r m5 diagnostics}

# Divergent transitions
np <- nuts_params(m5)
print(sum(subset(np, Parameter == "divergent__")$Value))

# \hat{R} ratio
print(max(brms::rhat(m5), na.rm = TRUE))

# ESS
min(neff_ratio(m5), na.rm = TRUE)
```

A disaster. Let's also plot the trace plots.

```{r m5 trace plot, message=FALSE}
traceplot(m5$fit)
```
Some chains are straight lines, and hence do not mix at all with chain #3!

We can also have a look at the distribution of `mutability`,
to realize how skewed it is.

```{r mutability plot}
(p.mutability <- (ggplot(by.statement, aes(x=mutability))
                  + geom_histogram(binwidth=1, fill="orange2")))
```

## Models $m_6$ and $m_7$

The most straightforward way out of this ditch
is to simply log-transform `mutability` (after adding 1 to all percentages
so that all logs are defined).

Furthermore, we ditch the multivariate model, since now we're
simply interested in estimating `einspect`. 

We also include another model $m_7$, which uses varying effects 
with grouping to model the interactions. Note that this kind of model
would have the same fitting issues as $m_5$ if we hadn't log-transformed
`mutability`.


```{r m6 m7 formula}
by.statement$logmutability <- log(1 + by.statement$mutability)

eq.m6 <- brmsformula(
  einspectS ~ 1 + (1|p|family) + (1|q|category) + 
    predicate*family + crashing*family + logmutability*family,
  family=brmsfamily("gaussian", link="log")
)

eq.m7 <- brmsformula(
  einspectS ~ 1 + (1|q|category) + (predicate + crashing + logmutability|p|family),
  family=brmsfamily("gaussian", link="log")
)

pp6.check <- get_prior(eq.m6, data=by.statement)
pp7.check <- get_prior(eq.m7, data=by.statement)

pp6 <- c(
  set_prior("normal(0, 1.0)", class="Intercept"),
  set_prior("normal(0, 1.0)", class="b"),
  set_prior("weibull(2, 0.3)", class="sd", group="family"),
  set_prior("weibull(2, 0.3)", class="sd", group="category"),
  set_prior("gamma(0.01, 0.01)", class="sigma")
)

pp7 <- c(
  set_prior("normal(0, 1.0)", class="Intercept"),
  set_prior("lkj(1)", class="cor"),
  set_prior("weibull(2, 0.3)", class="sd"),
  # set_prior("weibull(2, 0.3)", class="sd", group="category"),
  set_prior("gamma(0.01, 0.01)", class="sigma")
)
```

Alternative ways to make the models fittable include:

   - Introducing a multi-level term, with `einspect ~ log(x)*family`,
     and `log(x) = log(y) + a`, where $x/y = \textrm{mutability}$.
     This is based on rewriting $\log(a/b) = \alpha$ 
     into $\log(a) = \alpha + \log(b)$.
     
   - The approach followed in [this paper](https://arxiv.org/abs/2007.09394).

## Fitting $m_6$ and $m_7$

Prior checks:

```{r m6 priors check, warning=FALSE, message=FALSE, error=FALSE, results='hide'}
m6.priors <- brm(eq.m6,
                 prior=pp6,
                 data=by.statement,
                 chains=1, iter=1000,
                 sample_prior="only",
                 refresh=0, 
                 control=list(adapt_delta=0.95),
                 seed=MY.SEED)

(pp_check(m6.priors, draw_ids=1:100, resp="einspectS") + scale_x_log10())
```

```{r m7 priors check, warning=FALSE, message=FALSE, error=FALSE, results='hide'}
m7.priors <- brm(eq.m7,
                 prior=pp7,
                 data=by.statement,
                 chains=1, iter=1000,
                 sample_prior="only",
                 refresh=0, 
                 control=list(adapt_delta=0.95),
                 seed=MY.SEED)

(pp_check(m7.priors, draw_ids=1:100, resp="einspectS") + scale_x_log10())
```

We fit models $m_6$ and $m_7$.

```{r m6 m7 fit}
m6 <- brm(eq.m6, data=by.statement, prior=pp6, 
          refresh=0,
          control=list(adapt_delta=0.95), 
          seed=MY.SEED)
m7 <- brm(eq.m7, data=by.statement, prior=pp7, 
          refresh=0,
          control=list(adapt_delta=0.98), 
          seed=MY.SEED)
```

Diagnostics:

```{r m6 m7 diagnostics}

# Divergent transitions
np <- nuts_params(m6)
print(sum(subset(np, Parameter == "divergent__")$Value))

# \hat{R} ratio
print(max(brms::rhat(m6), na.rm = TRUE))

# ESS
min(neff_ratio(m6), na.rm = TRUE)

# Divergent transitions
np <- nuts_params(m7)
print(sum(subset(np, Parameter == "divergent__")$Value))

# \hat{R} ratio
print(max(brms::rhat(m7), na.rm = TRUE))

# ESS
min(neff_ratio(m7), na.rm = TRUE)
```

Posterior checks:

```{r m6 posterior checks, warning=FALSE, message=FALSE}
print(pp_check(m6, ndraws=100, resp="einspectS") + scale_x_log10())
```

```{r m7 posterior checks, warning=FALSE, message=FALSE}
print(pp_check(m7, ndraws=100, resp="einspectS") + scale_x_log10())
```

Everything is A-OK now. In fact, the posterior checks
look the best ones so far.

Let's compare the models using LOO.

```{r all model comparison, warning=FALSE}
m6 <- add_criterion(m6, criterion = "loo")
m7 <- add_criterion(m7, criterion = "loo")

(lc.m1m2m3m4m6m7 <- loo(m1, m2, m3, m4, m6, m7, moment_match = TRUE))
```

$m_6$ and $m_7$ are basically indistinguishable
in terms of predictive capabilities.
And, they both mop the floor with all other models (14 standard deviations of difference!). Note, however, that this is a result of
making them monovariate instead of multivariate;
if we kept them multi-variate, they would have predictive
capabilities close to $m_4$.


## Analyzing $m_6$ and $m_7$

```{r m6 summary}
summary(m6)
```

```{r m7 summary}
summary(m7)
```
Let's now analyze $m_6$ first, whose mutability-related effects
are fixed effects.

```{r m6 analysis}
MORE.PROBS <- c(0.5, 0.7, 0.8, 0.88, 0.9, 0.95, 0.99)
(res.m6.einspect.logmutability <- interval.results.interacts(
  m6, "einspectS", "logmutability",
  MORE.PROBS,
  bool=FALSE, multivariate=FALSE
  ))

# We can use m6's estimates for all these interactions,
# since they are very close to m4's
res.m6.einspect.crashing <- interval.results.interacts(
  m6, "einspectS", "crashing",
  MORE.PROBS,
  bool=TRUE, multivariate=FALSE
  )

res.m6.einspect.predicate <- interval.results.interacts(
  m6, "einspectS", "predicate",
  MORE.PROBS,
  bool=TRUE, multivariate=FALSE
  )
```
There is a tendency for MBFL to do better on mutable bugs, 
but it is not a strong tendency, and it can only be detected with 88%
confidence (which is still pretty clear).
Incidentally, PS and ST tend to perform *worse* on the same kinds of bugs,
whereas SBFL is agnostic.

Let's now switch to $m_7$ first, whose mutability-related effects
are varying effects.

```{r m7 analysis}
(res.m7.einspect.logmutability <- interval.results.ranef(m7, "einspectS", 
                                                         c("family",
                                                           "logmutability"),
                                                         MORE.PROBS,
                                                         multivariate=FALSE))
```
The results are kind of similar for MBFL, but now SBFL
also exhibits a similar tendency.
Altogether, we may prefer the simpler model $m_6$,
where the advantage of MBFL for mutable bugs
is not big but clearly discernible.

# Dump all constants

```{r storage}
store.results_ <- function(res)
{
  res_str <- str_split(deparse(substitute(res)), "[.]")[[1]]
  m <- res_str[2]
  outcome <- res_str[3]
  predictor <- res_str[4]
  ints <- res[["ints"]]
  est <- res[["est"]]
  cns <- colnames(ints)
  rws <- rownames(ints)
  percs <- round(100*as.numeric(str_extract(rws, "([0-9]+[.][0-9]+)")))
  tabname <- str_c(m, outcome, predictor, sep="/")
  store.table_(ints, tabname, replace.probs=TRUE)
  if (!is.null(est))
    store.table_(as.data.frame(est), tabname, replace.probs=TRUE, vectorize=TRUE)
}

store.results_(res.m1.einspect.family)
store.results_(res.m1.einspect.category)
store.results_(res.m1.time.family)
store.results_(res.m1.time.category)

store.results_(res.m2.einspect.family)
store.results_(res.m2.einspect.category)
store.results_(res.m2.time.family)
store.results_(res.m2.time.category)

store.results_(res.m3.einspect.family)

store.results_(res.m4.einspect.crashing)
store.results_(res.m4.einspect.predicate)
store.results_(res.m4.einspect.mutable)

store.results_(res.m6.einspect.logmutability)
store.results_(res.m6.einspect.crashing)
store.results_(res.m6.einspect.predicate)
```

```{r epilogue}
dump.all_(append=FALSE)
```